---
title: "Estimation of changes in air pollution during the COVID-19 outbreak"
author: "Group 3"
date: "`r Sys.Date()`"
output:
    html_document:
      number_sections: true
      highlight: haddock
      theme: spacelab
      toc: yes
      toc_depth: 2
      toc_float:
        collapsed: false
    
---

## Background:
The lockdown response to coronavirus disease 2019 (COVID-19) has caused an unprecedented decrease in global economic and transport activity. The movement of people and the corresponding activities of production and consumption have been significantly reduced. As a possible side effect of this reduction, air pollution in many areas may have been greatly reduced. Our group aim to estimate and visualize changes in air polluton during the COVID19.

## Objectives:
* We will look at the air pollution of four cities that have all been affected by COVID-19 at different levels. 
  From our initial discussion, we have selected the following: London (United Kingdom); Wuhan (China); New Delhi (India); Auckland (New Zealand).
  In details, we will look at changes in concentrations of NOx, NO2, PM, O3, VOC, NH3.
* We will look at the changes in the community mobility of these four cities before and after COVID-19.
* We aim to comparatively explore the changes in air pollution of these cities, comparing and contrasting the levels of ‘lockdown' measures put in place by local governments. 
* We will create informative visualisations to tell the COVID story, in addition to using ML techniques to understand the most significant factors of air pollution within the pandemic. Whilst we may have initial hypotheses, the data will speak for itself.

## Datasets:
The data we use came from [Air quality data](https://openaq.org/#/locations?_k=ggbbh8) and [COVID-19 Community Mobility Report Data](https://www.google.com/covid19/mobility/)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(ggridges)
library(gghalves)
library(lubridate)
library(gridExtra)
library(sf)
library(vroom)
library(skimr)
library(janitor)
library(mice)
library(patchwork)
```

# STEP 1 - EXPLORATORY ANALYSIS


```{r}
# first select mobility data for our target cities
in_mob <-  vroom("2020_IN_Region_Mobility_Report.csv") %>% filter(sub_region_1=="Delhi")

gb_mob <- vroom("2020_GB_Region_Mobility_Report.csv") %>% filter(sub_region_1=="Greater London")

ch_mob <- vroom("2020_CH_Region_Mobility_Report.csv") %>% filter(sub_region_1=="Zurich")
```

```{r}
# select only past 36 months
london_air <- vroom("london-air-quality.csv") %>% 
  mutate(date = ymd(date),
         year = year(date)) %>% filter(date>=as.Date("2018-01-01"))

delhi_air <- vroom("delhi-institute of tool engineering, wazirpur, delhi, delhi, india-air-quality.csv") %>% 
  mutate(date = ymd(date),
         year = year(date)) %>% filter(date>=as.Date("2018-01-01"))

wuhan_air <- vroom("wuhan-air-quality.csv") %>% 
  mutate(date = ymd(date),
         year = year(date)) %>% filter(date>=as.Date("2018-01-01"))

zurich_air <- vroom("zurich-kaserne, switzerland-air-quality.csv") %>% 
  mutate(date = ymd(date),
         year = year(date)) %>% filter(date>=as.Date("2018-01-01"))

```

```{r}
# describe the mobility data
skim(gb_mob)

skim(ch_mob)

skim(in_mob)
```

```{r}
# describe/view air quality data

skim(london_air)

skim(zurich_air)

skim(wuhan_air)

skim(delhi_air)

```

```{r}
# check for dupes
london_air%>%get_dupes(date)
wuhan_air%>%get_dupes(date)
zurich_air %>%get_dupes(date)
delhi_air %>% get_dupes(date)

# dupes in mobility?
in_mob %>% get_dupes(date)
ch_mob %>% get_dupes(date)
gb_mob %>% get_dupes(date)

# fix duplicates across sub regions by averaging the values
in_mob_clean <- in_mob %>% group_by(date) %>% 
  summarise(retail_and_recreation_percent_change_from_baseline=mean(retail_and_recreation_percent_change_from_baseline),
                                        grocery_and_pharmacy_percent_change_from_baseline=mean(grocery_and_pharmacy_percent_change_from_baseline),
                                        parks_percent_change_from_baseline=mean(parks_percent_change_from_baseline),
                                        transit_stations_percent_change_from_baseline=mean(transit_stations_percent_change_from_baseline),
                                        workplaces_percent_change_from_baseline=mean(workplaces_percent_change_from_baseline),
                                        residential_percent_change_from_baseline=mean(residential_percent_change_from_baseline)) %>% ungroup()
gb_mob_clean <- gb_mob %>% group_by(date) %>% 
  summarise(retail_and_recreation_percent_change_from_baseline=mean(retail_and_recreation_percent_change_from_baseline),
                                        grocery_and_pharmacy_percent_change_from_baseline=mean(grocery_and_pharmacy_percent_change_from_baseline),
                                        parks_percent_change_from_baseline=mean(parks_percent_change_from_baseline),
                                        transit_stations_percent_change_from_baseline=mean(transit_stations_percent_change_from_baseline),
                                        workplaces_percent_change_from_baseline=mean(workplaces_percent_change_from_baseline),
                                        residential_percent_change_from_baseline=mean(residential_percent_change_from_baseline)) %>% ungroup()

```
```{r}
# lets look into missing data
md.pattern(in_mob_clean,rotate.names = T)

md.pattern(gb_mob_clean,rotate.names = T)

md.pattern(ch_mob,rotate.names = T)
```

```{r}
# merge the mobility and air quality data
lon_merged <- left_join(london_air,gb_mob_clean,by='date')

del_merged <- left_join(delhi_air,in_mob_clean,by='date')

zur_merged <- left_join(zurich_air,ch_mob,by='date')
```

# Explorary Data Analysis

```{r fig.height=5,fig.width=12,warning=FALSE,message=FALSE}
# some visualizations
lon_air1 <- lon_merged %>% pivot_longer(cols=c(2:7),names_to="Pollutions",values_to="ugm")%>%
  filter(Pollutions %in% c("pm25","pm10","no2","so2")) %>%mutate(city = "London")
plt1<- ggplot(lon_air1) +
  geom_smooth(aes(x=date,y=ugm,color=Pollutions)) +
  labs(x="", y="Concentration (µg/m3)")+
  facet_wrap(~ city) + 
  theme_classic()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none",
         strip.text.x = element_text(),
        strip.text.y = element_text(),
        strip.background = element_rect(fill="#E1E1E1"))


del_air1 <- del_merged %>% pivot_longer(cols=c(2:7),names_to="Pollutions",values_to="ugm")%>%
  filter(Pollutions %in% c("pm25","pm10","no2","so2"))%>%mutate(city = "New Delhi")
plt2<- ggplot(del_air1) +
  geom_smooth(aes(x=date,y=ugm,color=Pollutions)) +
  labs(x="", y="Concentration (µg/m3)")+
  facet_wrap(~ city) + 
  theme_classic()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none",
         strip.text.x = element_text(),
        strip.text.y = element_text(),
        strip.background = element_rect(fill="#E1E1E1"))



zur_air1 <- zur_merged %>% pivot_longer(cols=c(2:7),names_to="Pollutions",values_to="ugm")%>%
  filter(Pollutions %in% c("pm25","pm10","no2","so2"))%>%mutate(city = "Zurich")
plt3<- ggplot(zur_air1) +
  geom_smooth(aes(x=date,y=ugm,color=Pollutions)) +
  labs(x="", y="Concentration (µg/m3)") +
  facet_wrap(~ city) + 
  theme_classic()+  
  theme(strip.text.x = element_text(size = 16)) +
   theme(
         strip.text.x = element_text(),
        strip.text.y = element_text()),
        strip.background = element_rect(fill="#E1E1E1"))


plt1+plt2+plt3+
  plot_annotation(title = "Air pollutant concentrations reached their lowest points in three years in July 2020",subtitle = "Air quality data of London, New Delhi and Zürich over the past three years",caption="Aqicn.org") &
    theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) 
    ) &
  theme(panel.background = element_rect(fill = NA,color="black"),               # Remove gridlines
        plot.title = element_text(,        # Change text format
                                  face = "bold", 
                                  size = 18),
        plot.subtitle = element_text(,
                                     size = 16),
        axis.title.x = element_blank(),
        axis.text.x = element_text(family = "Myanmar Text"),
        axis.title.y = element_text(family = "Myanmar Text"),
        axis.text.y = element_text(family = "Myanmar Text"),
        aspect.ratio = 3/5)

```
```{r fig.height=7,fig.width=12,warning=FALSE,message=FALSE}
lon_no2 <- london_air %>% summarise(date=date,London=no2)
del_no2 <- delhi_air %>% summarise(date=date,New_Delhi=no2)
zur_no2 <- zurich_air %>% summarise(date=date,Zurich=no2)
no2_merged <- left_join(lon_no2,del_no2,by='date') 
no2_merged <- left_join(no2_merged,zur_no2,by='date')%>% pivot_longer(cols=c(2:4),names_to="city",values_to="no2")%>% mutate(component = "NO2")

plt4 <- ggplot(no2_merged) + 
  geom_smooth(aes(x=date,y=no2,color=city)) +
  labs(x="", y="Concentration (µg/m3)")+
  facet_wrap(~ component) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
theme(legend.position = "none",
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

lon_pm10 <- london_air %>% summarise(date=date,London=pm10)
del_pm10 <- delhi_air %>% summarise(date=date,New_Delhi=pm10)
zur_pm10 <- zurich_air %>% summarise(date=date,Zurich=pm10)
pm10_merged <- left_join(lon_pm10,del_pm10,by='date') 
pm10_merged <- left_join(pm10_merged,zur_pm10,by='date')%>% pivot_longer(cols=c(2:4),names_to="city",values_to="pm10")%>% mutate(component = "PM10")

plt5 <- ggplot(pm10_merged) + 
  geom_smooth(aes(x=date,y=pm10,color=city)) +
  labs(x="", y="Concentration (µg/m3)")+
  facet_wrap(~ component) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme( 
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

lon_pm25 <- london_air %>% summarise(date=date,London=pm25)
del_pm25 <- delhi_air %>% summarise(date=date,New_Delhi=pm25)
zur_pm25 <- zurich_air %>% summarise(date=date,Zurich=pm25)
pm25_merged <- left_join(lon_pm25,del_pm25,by='date') 
pm25_merged <- left_join(pm25_merged,zur_pm25,by='date')%>% pivot_longer(cols=c(2:4),names_to="city",values_to="pm25")%>% mutate(component = "PM25")

plt6 <- ggplot(pm25_merged) + 
  geom_smooth(aes(x=date,y=pm25,color=city)) +
  labs(x="", y="Concentration (µg/m3)")+
  facet_wrap(~ component) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none",
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

lon_so2 <- london_air %>% summarise(date=date,London=so2)
del_so2 <- delhi_air %>% summarise(date=date,New_Delhi=so2)
zur_so2 <- zurich_air %>% summarise(date=date,Zurich=so2)
so2_merged <- left_join(lon_so2,del_so2,by='date') 
so2_merged <- left_join(so2_merged,zur_so2,by='date')%>% pivot_longer(cols=c(2:4),names_to="city",values_to="so2")%>% mutate(component = "SO2")

plt7 <- ggplot(so2_merged) + 
  geom_smooth(aes(x=date,y=so2,color=city)) +
  labs(x="", y="Concentration (µg/m3)")+
  facet_wrap(~ component) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

plt4+plt5+plt6+plt7+
  plot_annotation(title = "New Delhi has the most variable levels of pollutants throughout the year", subtitle = "Air quality data of London, New Delhi and Zürich over the past three years", caption="Aqicn.org") &
    theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) 
    ) &
      theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) 
    ) &
  theme(panel.background = element_rect(fill = NA,color="black"),               # Remove gridlines
        plot.title = element_text(family = "Myanmar Text",        # Change text format
                                  face = "bold", 
                                  size = 18),
        plot.subtitle = element_text(family = "Myanmar Text",
                                     size = 16),
        axis.title.x = element_blank(),
        axis.text.x = element_text(family = "Myanmar Text"),
        axis.title.y = element_text(family = "Myanmar Text"),
        axis.text.y = element_text(family = "Myanmar Text"),
        aspect.ratio = 3/5)

```

```{r fig.height=5,fig.width=12,warning=FALSE,message=FALSE}
# some visualizations
lon_pm25 <- lon_merged  %>% mutate(city ="London")
plta <- ggplot(lon_pm25,aes(x=date,y=pm25,color=year))+
  geom_smooth() +
  geom_point() +
  facet_wrap(~city)+
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none") +
  labs(x="")

del_pm25 <- del_merged %>% mutate(city="New Delhi")
pltb <- ggplot(del_pm25,aes(x=date,y=pm25,color=year)) +
  geom_smooth() +
  geom_point() +
  facet_wrap(~city)+
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none")+
  labs(x="")

zur_pm25 <- zur_merged %>% mutate(city="Zurich")
pltc <- ggplot(zur_pm25,aes(x=date,y=pm25,color=year)) +
  geom_smooth() +
  geom_point() +
  facet_wrap(~city)+
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  labs(x="")

plta+pltb+pltc+
  plot_annotation(title = " ", subtitle = "PM2.5 data of London, New Delhi and Zürich over the past three years", caption="Aqicn.org") &
    theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) 
    ) 
```

```{r fig.height=5,fig.width=12,warning=FALSE,message=FALSE}
lon_long <- lon_merged %>%
  pivot_longer(cols=c(9:14),names_to="type",values_to="percent") %>% filter(date>=as.Date("2020-02-15")) 
del_long <- del_merged %>% pivot_longer(cols=c(9:14),names_to="type",values_to="percent") %>% filter(date>=as.Date("2020-02-15"))
zur_long <- zur_merged %>% pivot_longer(cols=c(15:20),names_to="type",values_to="percent") %>% filter(date>=as.Date("2020-02-15"))

lon_mob <- lon_long %>% mutate (city = "London")

plt8 <- ggplot(lon_mob) + 
  geom_smooth(aes(x=date,y=percent,color=type)) +
  labs(x="", y="Percent change from baseline")+
  facet_wrap(~ city) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none") +
   theme(
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

del_mob <- del_long %>% mutate (city = "New Delhi")

plt9 <- ggplot(del_mob) + 
  geom_smooth(aes(x=date,y=percent,color=type)) +
  labs(x="", y="Percent change from baseline")+
  facet_wrap(~ city) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  theme(legend.position = "none") +
   theme(
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

zur_mob <- zur_merged %>%
  summarise(
    date = date,
    Retail = retail_and_recreation_percent_change_from_baseline,
    Grocery = grocery_and_pharmacy_percent_change_from_baseline,
    Parks = parks_percent_change_from_baseline,
    Transit = transit_stations_percent_change_from_baseline,
    Workplaces = workplaces_percent_change_from_baseline,
    Residential = residential_percent_change_from_baseline
  ) %>%pivot_longer(cols=c(2:7),names_to="type",values_to="percent") %>%filter(date>=as.Date("2020-02-15")) %>% mutate (city = "Zurich")

plt10 <- ggplot(zur_mob) + 
  geom_smooth(aes(x=date,y=percent,color=type)) +
  labs(x="", y="Percent change from baseline")+
  facet_wrap(~ city) + 
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
   theme(
         strip.text.x = element_text(family = "Myanmar Text"),
        strip.text.y = element_text(family = "Myanmar Text"),
        strip.background = element_rect(fill="#E1E1E1"))

plt8+plt9+plt10+
  plot_annotation(title = "Mobility Patterns Differed By City", subtitle = "Mobility quality data of London, New Delhi and Zürich since 15th Feberary", caption="Google") &

      theme(
  panel.background = element_rect(fill = NA,color="black"),               # Remove gridlines
        plot.title = element_text(family = "Myanmar Text",        # Change text format
                                  face = "bold", 
                                  size = 18),
        plot.subtitle = element_text(family = "Myanmar Text",
                                     size = 16),
        axis.title.x = element_blank(),
        axis.text.x = element_text(family = "Myanmar Text"),
        axis.title.y = element_text(family = "Myanmar Text"),
        axis.text.y = element_text(family = "Myanmar Text"),
        aspect.ratio = 3/5)


```

```{r fig.height = 10, fig.width = 12, warning = FALSE, message = FALSE}
# attempt one at percent change
del_pm2_averages <- del_merged %>% filter(year!=2020) %>% mutate(month=month(date),day=day(date)) %>% group_by(month,day) %>% summarise(base_pm25=mean(pm25))

del_long <- del_long %>% mutate(month=month(date),day=day(date)) %>%
  merge(.,del_pm2_averages,by=c("month","day")) %>%
  mutate(per_change_pm25=(pm25-base_pm25)/base_pm25*100)

lon_pm2_averages <- lon_merged %>% filter(year!=2020) %>% mutate(month=month(date),day=day(date)) %>% group_by(month,day) %>% summarise(base_pm25=mean(pm25))

lon_long <- lon_long %>% mutate(month=month(date),day=day(date)) %>%
  merge(.,lon_pm2_averages,by=c("month","day")) %>%
  mutate(per_change_pm25=(pm25-base_pm25)/base_pm25*100)

zur_pm2_averages <- zur_merged %>% filter(year!=2020) %>%
  mutate(month=month(date),day=day(date)) %>% group_by(month,day) %>% summarise(base_pm25=mean(pm25))

zur_long <- zur_long %>% mutate(month=month(date),day=day(date)) %>%
  merge(.,zur_pm2_averages,by=c("month","day")) %>%
  mutate(per_change_pm25=(pm25-base_pm25)/base_pm25*100)

lon_pm25_change <- lon_long %>% summarise(date=date,London=per_change_pm25)
del_pm25_change <- del_long %>% summarise(date=date,New_Delhi=per_change_pm25)
zur_pm25_change <- zur_long %>% summarise(date=date,Zurich=per_change_pm25)
pm25_change_merged <- left_join(lon_pm25_change,del_pm25_change,by='date') 
pm25_change_merged <- left_join(pm25_change_merged,zur_pm25_change,by='date')%>% pivot_longer(cols=c(2:4),names_to="city",values_to="per_change_pm25") %>% mutate(month=month(date))%>%
group_by(month,city)%>% summarise(mean_change_pm25 = mean(per_change_pm25))

plt11 <- ggplot(pm25_change_merged,aes(x=as.factor(month),y=mean_change_pm25)) +
  geom_bar(stat="identity",aes(color=city,fill=city))+
  labs(x="", y="Percent change of PM25 from baseline")+
  facet_wrap(~city) +
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  labs(title ="Percent change of PM2.5 from baseline in 2020", subtitle = "PM2.5 data of London, New Delhi and Zürich at 2020") +
  scale_x_discrete(breaks= c(2:12),labels= c("Feb", "Mar", "Apr","May", "Jun","July","Aug","Sep","Oct","Nov","Dec"))+
  theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) 
    ) +
    theme(legend.position = "none")


lon_pm25 <- lon_long %>% summarise(date=date,London=pm25)
del_pm25 <- del_long %>% summarise(date=date,New_Delhi=pm25)
zur_pm25 <- zur_long %>% summarise(date=date,Zurich=pm25)
pm25_merged <- left_join(lon_pm25,del_pm25,by='date') 
pm25_merged <- left_join(pm25_merged,zur_pm25,by='date')%>% pivot_longer(cols=c(2:4),names_to="city",values_to="pm25") %>% mutate(year = year(date),month=month(date))%>%filter(year==2020)%>%
group_by(year,month,city)%>% summarise(mean_pm25 = mean(pm25))

plt12 <- ggplot(pm25_merged,aes(x=as.factor(month),y=mean_pm25,label=round(mean_pm25,2))) +
  geom_bar(stat="identity",aes(color=city,fill=city))+
  geom_text(vjust = 0) + # add figures on each column
  labs(x="", y="Mean PM2.5 concentration")+
  coord_flip()+
  facet_wrap(~city,scales = "free") +
  theme_minimal()+
  theme(strip.text.x = element_text(size = 16)) +
  labs(title ="Mean PM2.5 concentration in 2020") +
  scale_x_discrete(breaks= c(2:12),labels= c("Feb", "Mar", "Apr","May", "Jun","July","Aug","Sep","Oct","Nov","Dec"))+
  theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) 
    ) +
    theme(legend.position = "none")


plt11/plt12 +
   
    theme(
    plot.title = element_text(face = "bold", size = 20, hjust = 0), 
  #hjust is to adjust the position of the title
    plot.subtitle = element_text(size = 16, hjust = 0) )

```
```{r}
# plot pm2 vs workplace percent change

del_long %>% 
  filter(date>=as.Date("2020-05-01")) %>% 
  ggplot(aes(y=per_change_pm25,x=percent)) +
  geom_smooth() + facet_wrap(~type,scales = "free")



lon_long %>% 
  filter(date>=as.Date("2020-05-01")) %>% 
  ggplot(aes(y=per_change_pm25,x=percent)) +
  geom_smooth() + facet_wrap(~type,scales = "free")


zur_long %>% 
  filter(date>=as.Date("2020-05-01")) %>% 
  ggplot(aes(y=per_change_pm25,x=percent)) +
  geom_smooth() + facet_wrap(~type,scales = "free")


```

```{r}
del_merged <- del_merged %>% mutate(month=month(date),day=day(date)) %>%
  merge(.,del_pm2_averages,by=c("month","day")) %>%
  mutate(per_change_pm25=(pm25-base_pm25)/base_pm25*100)
del_merged %>% 
  #filter(date>=as.Date("2020-01-01")) %>%
  ggplot(aes(y=per_change_pm25,x=date)) +
  geom_smooth()

lon_merged <- lon_merged %>% mutate(month=month(date),day=day(date)) %>%
  merge(.,del_pm2_averages,by=c("month","day")) %>%
  mutate(per_change_pm25=(pm25-base_pm25)/base_pm25*100)
lon_merged %>% 
  #filter(date>=as.Date("2020-01-01")) %>%
  ggplot(aes(y=per_change_pm25,x=date)) +
  geom_smooth()


zur_merged <- zur_merged %>% mutate(month=month(date),day=day(date)) %>%
  merge(.,del_pm2_averages,by=c("month","day")) %>%
  mutate(per_change_pm25=(pm25-base_pm25)/base_pm25*100)
zur_merged %>% 
  #filter(date>=as.Date("2020-01-01")) %>%
  ggplot(aes(y=per_change_pm25,x=date)) +
  geom_smooth()
```

# Regression

```{r warning = FALSE}

# CHANGES: no2 instead of pm25, and changed data set for delhi (was lon_merged now del_meged)

#zurich
m1 <- zur_merged %>% mutate(month=month(date)) %>% 
  lm(no2 ~ retail_and_recreation_percent_change_from_baseline+
     grocery_and_pharmacy_percent_change_from_baseline+
     transit_stations_percent_change_from_baseline+
     parks_percent_change_from_baseline+
     workplaces_percent_change_from_baseline+
     residential_percent_change_from_baseline+
       month,na.action=na.omit,data=.)

summary(m1) # Adjusted R2: 0.1808
# Most significant parameters: intercept , residential , month

#london
#m2 <- lon_merged %>% mutate(month=month(date)) %>% 
#  lm(no2 ~ retail_and_recreation_percent_change_from_baseline+
#     grocery_and_pharmacy_percent_change_from_baseline+
#     transit_stations_percent_change_from_baseline+
#     parks_percent_change_from_baseline+
#     workplaces_percent_change_from_baseline+
#     residential_percent_change_from_baseline+
#       month,na.action=na.omit,data=.)

#summary(m2) # Adjusted R2: 0.5793
# Most significant parameters: intercept - only other significant param is grocery_and_pharmacy

#delhi
m3 <- del_merged %>% mutate(month=month(date)) %>% 
  lm(no2 ~ retail_and_recreation_percent_change_from_baseline+
     grocery_and_pharmacy_percent_change_from_baseline+
     transit_stations_percent_change_from_baseline+
     parks_percent_change_from_baseline+
     workplaces_percent_change_from_baseline+
     residential_percent_change_from_baseline+
       month,na.action=na.omit,data=.)

summary(m3) # Adjusted R2: 0.4941
# Most significant parameters: intercept , retail_and_recreation , parks , month

```

# Random Forest
```{r, rf_delhi,fig.height=5,fig.width=8}

# CHANGED: whole chunk
library(caret)
library(rpart.plot)

train_control <- trainControl (
    method="cv",
    number=5,
    verboseIter=TRUE)

library(GGally)

set.seed(1234)
# Fit random forest: model
merged_nona <- del_merged %>% mutate(month=month(date))
merged_nona %>% ggcorr(layout.position="left")
rf_RF <- train(
   no2 ~ retail_and_recreation_percent_change_from_baseline+
     grocery_and_pharmacy_percent_change_from_baseline+
     transit_stations_percent_change_from_baseline+
     retail_and_recreation_percent_change_from_baseline+
     workplaces_percent_change_from_baseline+
     residential_percent_change_from_baseline+
    month,
  merged_nona, 
  method = "ranger",
  na.action = na.omit,
   #metric="ROC",
  trControl = train_control,
  importance = 'permutation',
  tuneLength=10
)

rf_RF$results

rf_RF$bestTune

importance <- varImp(rf_RF, scale=TRUE)
importance
varImps <- importance$importance %>% rownames_to_column()
# Plot variable inportance bar chart
varImps %>% ggplot(aes(x = reorder(rowname,-Overall),
                       y = Overall)) +
  geom_col(fill = "#74AEDB",
           color = "#4F96CD") +
  scale_x_discrete(labels=c("Month","Retail & Recreation","Transit Stations",
                            "Residential","Grocery & Pharmacy","Workplaces"))+
  theme_classic() +
  labs(title = "New Delhi's air quality was predominantly influenced by the time of the year rather than 
changes in mobility",
       subtitle = "Feature importance of mobility categories and month towards levels of NO2 in 2020",
       y = "Feature importance weight", 
       x = "") +
  theme(panel.background = element_rect(fill = NA),               # Remove gridlines
        plot.title = element_text(        # Change text format
                                  face = "bold", 
                                  size = 12),
        plot.subtitle = element_text(
                                     size = 10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(),
        axis.title.y = element_text(),
        axis.text.y = element_text(),
        aspect.ratio = 3/5)

```



```{r rf_london}

# CHANGED: whole chunk

train_control <- trainControl (
    method="cv",
    number=5,
    verboseIter=TRUE)

merged_nona %>% ggcorr(layout.position="left")
set.seed(1234)
# Fit random forest: model
merged_nona <- lon_merged %>% mutate(month=month(date))
rf_RF <- train(
   no2 ~ retail_and_recreation_percent_change_from_baseline+
     grocery_and_pharmacy_percent_change_from_baseline+
     transit_stations_percent_change_from_baseline+
     retail_and_recreation_percent_change_from_baseline+
     workplaces_percent_change_from_baseline+
     residential_percent_change_from_baseline+
    month,
  merged_nona, 
  method = "ranger",
  na.action = na.omit,
   #metric="ROC",
  trControl = train_control,
  importance = 'permutation',
  tuneLength=10
)

rf_RF$results

rf_RF$bestTune

importance <- varImp(rf_RF, scale=TRUE)
importance
varImps <- importance$importance %>% rownames_to_column()
# Plot variable inportance bar chart
varImps %>% ggplot(aes(x = reorder(rowname,-Overall),
                       y = Overall)) +
  geom_col(fill = "#74AEDB",
           color = "#4F96CD") +
  scale_x_discrete(labels=c("Workplaces", "Transit Stations", "Retail & Recreation", "Residential", "Month", "Grocery & Pharmacy")) +
  theme_classic() +
  labs(title = "London's air quality was highly affected by mobility patterns rather than the month",
       subtitle = "Feature importance of mobility categories and month towards levels of NO2 in 2020",
       y = "Feature importance weight", 
       x = "") +
  theme(panel.background = element_rect(fill = NA),               # Remove gridlines
        plot.title = element_text(,        # Change text format
                                  face = "bold", 
                                  size = 12),
        plot.subtitle = element_text(,
                                     size = 10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(),
        axis.title.y = element_text(),
        axis.text.y = element_text())

```
```{r, rf_zurich}

# CHANGED: whole chunk

train_control <- trainControl (
    method="cv",
    number=5,
    verboseIter=TRUE)

merged_nona %>% ggcorr(layout.position="left")
set.seed(1234)
# Fit random forest: model
merged_nona <- zur_merged %>% mutate(month=month(date))
rf_RF <- train(
   no2 ~ retail_and_recreation_percent_change_from_baseline +
     grocery_and_pharmacy_percent_change_from_baseline +
     transit_stations_percent_change_from_baseline +
     retail_and_recreation_percent_change_from_baseline +
     workplaces_percent_change_from_baseline +
     residential_percent_change_from_baseline +
    month,
  merged_nona, 
  method = "ranger",
  na.action = na.omit,
   #metric="ROC",
  trControl = train_control,
  importance = 'permutation',
  tuneLength=10
)

rf_RF$results

rf_RF$bestTune

importance <- varImp(rf_RF, scale=TRUE)
importance
varImps <- importance$importance %>% rownames_to_column()
# Plot variable inportance bar chart
varImps %>% ggplot(aes(x = reorder(rowname,-Overall),
                       y = Overall)) +
  geom_col(fill = "#74AEDB",
           color = "#4F96CD") +
  scale_x_discrete(labels=c("Month","Retail & Recreation", "Transit Stations", "Residential",  "Workplaces", "Grocery & Pharmacy")) +
  theme_classic() +
  labs(title = "Zurich's air quality stayed largely unaffected by changes in mobility",
       subtitle = "Feature importance of mobility categories and month towards levels of NO2 in 2020",
       y = "Feature importance weight", 
       x = "") +
  theme(panel.background = element_rect(fill = NA),               # Remove gridlines
        plot.title = element_text(,        # Change text format
                                  face = "bold", 
                                  size = 12),
        plot.subtitle = element_text(,
                                     size = 10),
        axis.title.x = element_blank(),
        axis.text.x = element_text(),
        axis.title.y = element_text(),
        axis.text.y = element_text())

```



